---
title: 'Practical Machine Learning - Submission'
output: html_document
---

##Objective
The goal of this work is to predict from data collected with quantified self devices if people are performing barbell lifts correctly or incorrectly.


##Librairies & seed
```{r, echo=TRUE}
library(caret)
set.seed(1234)
```

##Collect data and build data sets
```{r, echo=TRUE}
url_csv_train="http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
url_csv_test="http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

raw_data_train <- read.csv(url_csv_train)
raw_data_test <- read.csv(url_csv_test)


inTrain <- createDataPartition(raw_data_train$classe, p = 0.7, list = FALSE)
training <- raw_data_train[ inTrain,]
testing <- raw_data_train[-inTrain,]
```

##Features
By looking at the data (using str), we can see that a majority of variables are filled with a large majority of NA, or are empty.
We decide to eliminate:
- these variables with a large majority of NA or empty values
- the first columns (X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, new_window, num_window) which may not give information on how well the exercise is performed

```{r, echo=TRUE}
features=c("roll_belt","pitch_belt","yaw_belt","total_accel_belt","gyros_belt_x","gyros_belt_y","gyros_belt_z","accel_belt_x","accel_belt_y","accel_belt_z","magnet_belt_x","magnet_belt_y","magnet_belt_z","roll_arm","pitch_arm","yaw_arm","total_accel_arm","gyros_arm_x","gyros_arm_y","gyros_arm_z","accel_arm_x","accel_arm_y","accel_arm_z","magnet_arm_x","magnet_arm_y","magnet_arm_z","roll_dumbbell","pitch_dumbbell","yaw_dumbbell","total_accel_dumbbell","gyros_dumbbell_x","gyros_dumbbell_y","gyros_dumbbell_z","accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z","magnet_dumbbell_x","magnet_dumbbell_y","magnet_dumbbell_z","roll_forearm","pitch_forearm","yaw_forearm","total_accel_forearm","gyros_forearm_x","gyros_forearm_y","gyros_forearm_z","accel_forearm_x","accel_forearm_y","accel_forearm_z","magnet_forearm_x","magnet_forearm_y","magnet_forearm_z")
```

At the end, we keep the variables which seems to be the measures:
- roll
- pitch
- yaw
- total accel
- gyros x
- gyros y 
- gyros z
- accel x
- accel y
- accel z
- magnet x
- magnet y
- magnet z

for different parts/captors:
- belt
- arm
- dumbbell
- forearm


##Model
We choose "randomForest" algorithm
and to perform 5-fold cross-validation.

```{r, echo=TRUE}
control1 <- trainControl(method = "cv", number = 5, allowParallel = TRUE)
modelFit <- train(classe~.,data=training[,append(features,"classe")],trControl = control1,method="rf")
```

##Test - out of sample error
We use our own testing data set to determine the out of sample error.
```{r, echo=TRUE}
predictions<-predict(modelFit,newdata=testing)
confusionMatrix(predictions,testing$classe)
```

##Predict the values for the test set
```{r, echo=TRUE}
answers=predict(modelFit,newdata=raw_data_test)
answers


pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(answers)
```
